### 2020-03-18 text/manuscript.tex
Add draft introduction

% Template for PLoS
% Version 3.5 March 2018
%
% % % % % % % % % % % % % % % % % % % % % %
%
% -- IMPORTANT NOTE
%
% This template contains comments intended 
% to minimize problems and delays during our production 
% process. Please follow the template instructions
% whenever possible.
%
% % % % % % % % % % % % % % % % % % % % % % % 
%
% Once your paper is accepted for publication, 
% PLEASE REMOVE ALL TRACKED CHANGES in this file 
% and leave only the final text of your manuscript. 
% PLOS recommends the use of latexdiff to track changes during review, as this will help to maintain a clean tex file.
% Visit https://www.ctan.org/pkg/latexdiff?lang=en for info or contact us at latex@plos.org.
%
%
% There are no restrictions on package use within the LaTeX files except that 
% no packages listed in the template may be deleted.
%
% Please do not include colors or graphics in the text.
%
% The manuscript LaTeX source should be contained within a single file (do not use \input, \externaldocument, or similar commands).
%
% % % % % % % % % % % % % % % % % % % % % % %
%
% -- FIGURES AND TABLES
%
% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
%
% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
% - Figures should be uploaded separately from your manuscript file. 
% - Figures generated using LaTeX should be extracted and removed from the PDF before submission. 
% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
% For figure citations, please use "Fig" instead of "Figure".
% See http://journals.plos.org/plosone/s/figures for PLOS figure guidelines.
%
% Tables should be cell-based and may not contain:
% - spacing/line breaks within cells to alter layout or alignment
% - do not nest tabular environments (no tabular environments within tabular environments)
% - no graphics or colored text (cell background color/shading OK)
% See http://journals.plos.org/plosone/s/tables for table guidelines.
%
% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
%
% % % % % % % % % % % % % % % % % % % % % % % %
%
% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
%
% IMPORTANT
% Below are a few tips to help format your equations and other special characters according to our specifications. For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://journals.plos.org/plosone/s/latex
%
% For inline equations, please be sure to include all portions of an equation in the math environment.  For example, x$^2$ is incorrect; this should be formatted as $x^2$ (or $\mathrm{x}^2$ if the romanized font is desired).
%
% Do not include text that is not math in the math environment. For example, CO2 should be written as CO\textsubscript{2} instead of CO$_2$.
%
% Please add line breaks to long display equations when possible in order to fit size of the column. 
%
% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
%
% When adding superscript or subscripts outside of brackets/braces, please group using {}.  For example, change "[U(D,E,\gamma)]^2" to "{[U(D,E,\gamma)]}^2". 
%
% Do not use \cal for caligraphic font.  Instead, use \mathcal{}
%
% % % % % % % % % % % % % % % % % % % % % % % % 
%
% Please contact latex@plos.org with any questions.
%
% % % % % % % % % % % % % % % % % % % % % % % %

\documentclass[10pt,letterpaper]{article}
\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}

% amsmath and amssymb packages, useful for mathematical formulas and symbols
\usepackage{amsmath,amssymb}

% Use adjustwidth environment to exceed column width (see example table in text)
\usepackage{changepage}

% Use Unicode characters when possible
\usepackage[utf8x]{inputenc}

% textcomp package and marvosym package for additional characters
\usepackage{textcomp,marvosym}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

% Use nameref to cite supporting information files (see Supporting Information section for more info)
\usepackage{nameref,hyperref}

% line numbers
\usepackage[right]{lineno}

% ligatures disabled
\usepackage{microtype}
\DisableLigatures[f]{encoding = *, family = * }

% color can be used to apply background shading to table cells only
\usepackage[table]{xcolor}

% array package and thick rules for tables
\usepackage{array}

% create "+" rule type for thick vertical lines
\newcolumntype{+}{!{\vrule width 2pt}}

% create \thickcline for thick horizontal lines of variable length
\newlength\savedwidth
\newcommand\thickcline[1]{%
  \noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
  \cline{#1}%
  \noalign{\vskip\arrayrulewidth}%
  \noalign{\global\arrayrulewidth\savedwidth}%
}

% \thickhline command for thick horizontal lines that span the table
\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
\hline
\noalign{\global\arrayrulewidth\savedwidth}}

% Remove comment for double spacing
%\usepackage{setspace} 
%\doublespacing

% Text layout
\raggedright
\setlength{\parindent}{0.5cm}
\textwidth 5.25in 
\textheight 8.75in

% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
% Captions will be left justified
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
\renewcommand{\figurename}{Fig}

% Use the PLoS provided BiBTeX style
\bibliographystyle{plos2015}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Header and Footer with logo
\usepackage{lastpage,fancyhdr,graphicx}
\usepackage{epstopdf}
%\pagestyle{myheadings}
\pagestyle{fancy}
\fancyhf{}
%\setlength{\headheight}{27.023pt}
%\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
\fancyheadoffset[L]{2.25in}
\fancyfootoffset[L]{2.25in}
\lfoot{\today}

%% Include all macros below

\newcommand{\lorem}{{\bf LOREM}}
\newcommand{\ipsum}{{\bf IPSUM}}

%% END MACROS SECTION

\begin{document}
\vspace*{0.2in}

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Which sectors are most prone to social distancing measures?} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Miklós Koren\textsuperscript{1,2,3\Yinyang\textcurrency*},
Rita Pető\textsuperscript{2\ddag}
\\
\bigskip
\textbf{1} Central European University, Budapest, Hungary
\\
\textbf{2} KRTK KTI, Budapest, Hungary
\\
\textbf{3} CEPR, London, UK
\\
\bigskip

% Insert additional author notes using the symbols described below. Insert symbol callouts after author names as necessary.
% 
% Remove or comment out the author notes below if they aren't used.
%
% Primary Equal Contribution Note
\Yinyang This author is responsible for the research idea, the theoretical model, the analysis of the spatial data and the first draft of the manuscript.

% Additional Equal Contribution Note
% Also use this double-dagger symbol for special authorship notes, such as senior authorship.
\ddag This author is responsible for the analysis of the occupation data, reviewing and discussing results and editing the manuscript.

% Current address notes
\textcurrency Current Address: Central European University, Nádor u. 9., 1051 Budapest, Hungary % change symbol to "\textcurrency a" if more than one current address note
% \textcurrency b Insert second current address 
% \textcurrency c Insert third current address

% Use the asterisk to denote corresponding authorship and provide email address in note below.
* korenm@ceu.edu

\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
Social distancing interventions can be effective against epidemics, but little is known about their economic effects. We develop a model of communication and production to study the effects of a policy that limits face-to-face interactions. This policy results in the highest cost increases in sectors with high division of labor and hence high need for communication and those that are located in urban centers with high population density. We measure the prevalence of communication-related tasks and the urban density of workers for each detailed industry in the U.S. We predict that XX, XX and XX will suffer the biggest employment losses in response to the 2019--20 coronavirus pandemic. Overall, XX people work in the most affected sectors.

\linenumbers

% Use "Eq" instead of "Equation" for equation citations.
\section*{Introduction}
Social distancing measures are effective non-pharmaceutical interventions against the rapid spread of epidemics (REFs). Many countries have implemented or are considering measures such as school closures, prohibition of large gatherings and transportation to slow down the spread of the coronavirus (REFs). What are the economic effects of such social distancing interventions? Which are the firms and industries most affected by the restrictions?

Past research has analyzed the efficacy of social distancing interventions on reducing the spread of epidemics using the 1918 Spanish Flu in the U.S. (REF) and seasonal viral infections in France (REF). Our knowledge of economic impacts, however, is limited. For this question, the case of the Spanish Flu pandemic is less relevant, as the importance of face-to-face communication has increased in the last 100 years through urbanization (REF) and an increased specialization in business services (REF).

We build a model of communication within the production process to understand how limiting face-to-face interaction increases costs. Without social distancing, workers specialize in a narrow range tasks and interact with other workers completing other tasks. This division of labor reduces production costs but requires frequent contact between workers. In the model, the number of contacts per worker are the most frequent in high population density areas in sectors where division of labor is important. When face-to-face interaction is limited, these are the locations and sectors that suffer the largest cost increases. Our model can also explain when telecommunication is a low-cost substitute for face-to-face interaction.

To measure sensitivity to social distancing, we turn to recent data on the task descriptions of each occupation (REF) and the precise geographic location of non-farm businesses (REF). We classify workers as affected if their occupation requires frequent communication with others (inside or outside the firm). The most affected group of workers includes healthcare professionals, teachers, and managers. Sectors are affected if they employ most of their workers in high-density ZIP-code areas. The most affected sectors are apparel manufacturing, information and business services, and performing arts. Both measures are necessary to understand exposure to social distancing. Firms may also locate in urban centers to minimize shipping costs with respect to suppliers and customers. But if they also have a high prevalence of communication-related tasks, we conclude that they are heavily relying on face-to-face communication, and will be particularly hurt by social distancing.

For completeness, we also compute, for each industry, the fraction of workers who work with many coworkers in enclosed space or subject to potential infections, or working in large plants. These additional measures are highly correlated with our two preferred indicators (CHECK). We report our four index of exposure to social distancing for each 5-digit occupation, each 3-digit NAICS industry, and each 5-digit ZIP-Code Tabulation Area. Industries that are in the top quartile of both  communication exposure and urban density altogether employ XX workers in the U.S.
\section{A model of communication}

Production involves sequentally completing tasks indexed by $z\in[0,1]$. A single worker can do a range of tasks, but there is benefit to specialization and division of labor (Adam Smith and Acemogul> REF). The labor cost of a worker completing $Z<1$ measure of tasks is $wZ^{1+\gamma}/\gamma$, where $w$ is the wage rate, and $\gamma>0$ captures the benefits to division of labor. As we show below, the higher the $\gamma$, the more specialized each worker will be in a narrower set of tasks.

Once the range of tasks $Z$ is completed, the worker passes the unfinished product on to another worker. This has a cost $c$, which can capture the cost of communication either face-to-face or online. The determinants of communication cost will parametrized later.

The firm will optimally decide how to share tasks between workers. The key trade-off is economizing on the cost of communication while exploiting the division of labor. Let $n$ denote the number of workers involved in the production process. Because workers are symmetric, each works on $Z=1/n$ range of tasks before passing the work to the next worker. The process involves $n-1$ ``contacts'' (instances of communication), which can be approximated by $n$ if $n$ is sufficiently large.

The firm's cost minimization problem can then be written as a function of the number of contacts alone,
\begin{equation}
	\min_n nc + \frac w\gamma n^{-\gamma},
\end{equation}
where total communication costs are $nc$ and production costs are $w n Z^{1+\gamma}/\gamma$ with $Z=1/n$.

Given the strict convexity of this cost function, the first-order condition is necessary and sufficient for the optimum,
\begin{equation}
	n^*(c) = (c/w)^{-1/(1+\gamma)}.
\end{equation}
The number of worker contacts is decreasing in the cost of communication, expressed relative to worker wage. When the division of labor is important, $\gamma$ is high, and the number of contacts does not depend very strongly on communication costs.

The total cost of producing one good can be calculated by substituting in \eqref{} into \eqref{},
\begin{equation}
	\frac1\chi c^{\chi}w^{1-\chi},
\end{equation}
where $\chi=\gamma/(1+\gamma)\in(0,1)$ measures the importance of division of labor. This unit cost function is the same as if workers and communication were substitutable in the production function in a Cobb-Douglas fashion. Here $\chi$ captures the share of costs associated with communication. [NB: and could hence be calibrated to the share of non-production workers]
\subsection{The cost of communication}
Firms can chose between two ways of organizing communication. First, they can ask workers to meet face to face, in which case communication costs depends inversely on the population density in the neighborhood of the firm, $c=d^{-\tau}$. This captures the Marshallian externality of knowledge spillovers (REF), which happen more easily in densely populated areas. Second, they can ask workers to  interact with one another through telecommunication, in which case $c=T$. In sectors where telecommunication is not feasible, we can take $T$ to be prohibitively large.

The firm will chose the cheaper of the two forms of communcation,
\begin{equation}
	c = \min \{d^{-\tau}, T\}.
\end{equation}
The unit cost of production can be calculated by substituting in \eqref{} into \eqref{},
\begin{equation}
\frac {w^{1-\chi}}{\chi}
\min \{
	d^{-\tau\chi},T^\chi
\}.
\end{equation}
Because both wages and land rents are higher in places with high population density (REF), a firm will only move to such location when either the communication benefits $\tau\chi$ are high, or the cost of telecommunication $T$ is high. In spatial equilibrium (not modeled here), firms with high communication needs will choose to locate in high-density areas.

\subsection{Social distancing measures}
We study the effect of a of social distancing intervention that introduces an upper bound $N$ on the number of face-to-face contacts, but places no restriction on telecommunication.

Take a firm that is using face-to-face communication, such that $d>T^{-1/\tau}$. (If the firm is already relying on telecommunication, they are unaffected by social distancing.) The optimal number of contacts without social distancing would be $n^* = d^{\tau(1-\chi)}w^{1-\chi}$. The firm is affected by the first type of social distancing if $n^*>N$, that is, 
\begin{equation}
	d > N^{1/\tau(1-\chi)} w^{-1/\tau}.
\end{equation}
This is more likely to hold for firms in high-density areas (high $d$) and those with high division of labor ($\chi$ close to one). These are both contributors to a high number of face-to-face contacts. Density and the prevalence of communication are positive predictors to sensitivity to social distancing.

How do these firms react to social distancing? They can limit face-to-face communication and decrease worker productivity (because each worker has to work on a wider range of tasks than they used to). [NB: cite digital education as an example in the introduction] Or they can mandate telecommuting and avoid face-to-face communication this way. The firm will chose the strategy that leads to the least increase in costs. They will switch to telecommuting if
\begin{equation}
	T < w N^{\chi-1}.
\end{equation}
Firms with low telecommunication costs $T$, high worker wages $w$ and low communication intensity $\chi$ are more likely to switch to telecommuting. Telecommuting is also more likely for more stringent social distancing measures (low $N$). Firms with high communication intensity are hurt the most by social distancing, because for them telecommuting is not a viable option.

\subsection{Discussion}

\section{Methodology}
We compute two measures of exposure to social distancing interventions. The first is the share of workers whose job requires personal communications. This is high for sectors like XX and low for sectors like XX. The idea is that when workers in communication-intensive occupations stay at home, their productivity decreases. A challenge to this assumption is that, in some occupations, online communication can partly substitute face-to-face communication.

The second measure captures how urbanized workers are: what fraction of them work in high population or employment density areas. This is high for sectors like XX and low for sectors like XX. When the demand for face-to-face communication, either within the firm or with consumers is large, firms are likely to agglomerate in large cities. The challenge to this measure is that agglomeration may also be driven by a minimization of transportation costs.

Taken together, high values of both measures of communication intensity and agglomeration coefficient can indicate that face-to-face communication (as opposed to online communication or transportation) is used intensively in the sector.

\section{Data}
ONET
CBP density
CBP plant size

Cross walks
Construction of communication index
Worker density
% Place figure captions after the first paragraph in which they are cited.
\begin{figure}[!h]
\caption{{\bf Bold the figure title.}
Figure caption text here, please use this space for the figure panel descriptions instead of using subfigure commands. A: Lorem ipsum dolor sit amet. B: Consectetur adipiscing elit.}
\label{fig1}
\end{figure}

% Results and Discussion can be combined.
\section*{Results}

% Place tables after the first paragraph in which they are cited.
\begin{table}[!ht]
\begin{adjustwidth}{-2.25in}{0in} % Comment out/remove adjustwidth environment if table fits in text column.
\centering
\caption{
{\bf Table caption Nulla mi mi, venenatis sed ipsum varius, volutpat euismod diam.}}
\begin{tabular}{|l+l|l|l|l|l|l|l|}
\hline
\multicolumn{4}{|l|}{\bf Heading1} & \multicolumn{4}{|l|}{\bf Heading2}\\ \thickhline
$cell1 row1$ & cell2 row 1 & cell3 row 1 & cell4 row 1 & cell5 row 1 & cell6 row 1 & cell7 row 1 & cell8 row 1\\ \hline
$cell1 row2$ & cell2 row 2 & cell3 row 2 & cell4 row 2 & cell5 row 2 & cell6 row 2 & cell7 row 2 & cell8 row 2\\ \hline
$cell1 row3$ & cell2 row 3 & cell3 row 3 & cell4 row 3 & cell5 row 3 & cell6 row 3 & cell7 row 3 & cell8 row 3\\ \hline
\end{tabular}
\begin{flushleft} Table notes Phasellus venenatis, tortor nec vestibulum mattis, massa tortor interdum felis, nec pellentesque metus tortor nec nisl. Ut ornare mauris tellus, vel dapibus arcu suscipit sed.
\end{flushleft}
\label{table1}
\end{adjustwidth}
\end{table}

\section*{Discussion}
Nulla mi mi, venenatis sed ipsum varius, Table~\ref{table1} volutpat euismod diam. Proin rutrum vel massa non gravida. Quisque tempor sem et dignissim rutrum. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Morbi at justo vitae nulla elementum commodo eu id massa. In vitae diam ac augue semper tincidunt eu ut eros. Fusce fringilla erat porttitor lectus cursus, vel sagittis arcu lobortis. Aliquam in enim semper, aliquam massa id, cursus neque. Praesent faucibus semper libero~\cite{bib3}.

\section*{Conclusion}

CO\textsubscript{2} Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque. Quisque augue sem, tincidunt sit amet feugiat eget, ullamcorper sed velit.

Sed non aliquet felis. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Mauris commodo justo ac dui pretium imperdiet. Sed suscipit iaculis mi at feugiat. Ut neque ipsum, luctus id lacus ut, laoreet scelerisque urna. Phasellus venenatis, tortor nec vestibulum mattis, massa tortor interdum felis, nec pellentesque metus tortor nec nisl. Ut ornare mauris tellus, vel dapibus arcu suscipit sed. Nam condimentum sem eget mollis euismod. Nullam dui urna, gravida venenatis dui et, tincidunt sodales ex. Nunc est dui, sodales sed mauris nec, auctor sagittis leo. Aliquam tincidunt, ex in facilisis elementum, libero lectus luctus est, non vulputate nisl augue at dolor. For more information, see \nameref{S1_Appendix}.

\section*{Supporting information}

% Include only the SI item label in the paragraph heading. Use the \nameref{label} command to cite SI items in the text.
\paragraph*{S1 Fig.}
\label{S1_Fig}
{\bf Bold the title sentence.} Add descriptive text after the title of the item (optional).

\paragraph*{S2 Fig.}
\label{S2_Fig}
{\bf Lorem ipsum.} Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque.

\paragraph*{S1 File.}
\label{S1_File}
{\bf Lorem ipsum.}  Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque.

\paragraph*{S1 Video.}
\label{S1_Video}
{\bf Lorem ipsum.}  Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque.

\paragraph*{S1 Appendix.}
\label{S1_Appendix}
{\bf Lorem ipsum.} Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque.

\paragraph*{S1 Table.}
\label{S1_Table}
{\bf Lorem ipsum.} Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque.

\section*{Acknowledgments}
Cras egestas velit mauris, eu mollis turpis pellentesque sit amet. Interdum et malesuada fames ac ante ipsum primis in faucibus. Nam id pretium nisi. Sed ac quam id nisi malesuada congue. Sed interdum aliquet augue, at pellentesque quam rhoncus vitae.

\nolinenumbers

% Either type in your references using
% \begin{thebibliography}{}
% \bibitem{}
% Text
% \end{thebibliography}
%
% or
%
% Compile your BiBTeX database using our plos2015.bst
% style file and paste the contents of your .bbl file
% here. See http://journals.plos.org/plosone/s/latex for 
% step-by-step instructions.
% 
\begin{thebibliography}{10}

\bibitem{bib1}
Conant GC, Wolfe KH.
\newblock {{T}urning a hobby into a job: how duplicated genes find new
  functions}.
\newblock Nat Rev Genet. 2008 Dec;9(12):938--950.

\bibitem{bib2}
Ohno S.
\newblock Evolution by gene duplication.
\newblock London: George Alien \& Unwin Ltd. Berlin, Heidelberg and New York:
  Springer-Verlag.; 1970.

\bibitem{bib3}
Magwire MM, Bayer F, Webster CL, Cao C, Jiggins FM.
\newblock {{S}uccessive increases in the resistance of {D}rosophila to viral
  infection through a transposon insertion followed by a {D}uplication}.
\newblock PLoS Genet. 2011 Oct;7(10):e1002337.

\end{thebibliography}

\end{document}