### 2020-06-30 text/regression.tex
Reformat regression table to fit in PLOS column width
Resolves #9

{\setlength{\tabcolsep}{0pt}
{\def\sym#1{\ifmmode^{#1}\else\(^{#1}\)\fi}\begin{tabular}{l*{4}{D{.}{.}{-1}}}\hline\hline
                    &\multicolumn{1}{c}{(1)}&\multicolumn{1}{c}{(2)}&\multicolumn{1}{c}{(3)}&\multicolumn{1}{c}{(4)}\\
                    &\multicolumn{1}{c}{} &\multicolumn{1}{c}{} &\multicolumn{1}{c}{} &\multicolumn{1}{c}{} \\
\hline
Customer-facing~workers&      -0.293\sym{**} &                     &                     &                     \\
(share, [0,1])      &     (0.138)         &                     &                     &                     \\
[1em]
--of~which~face-to-face&                     &      -0.409\sym{**} &      -0.453\sym{***}&       0.066         \\
intensive           &                     &     (0.158)         &     (0.142)         &     (0.155)         \\
[1em]
Teamwork-intensive~workers&       0.677\sym{***}&                     &                     &                     \\
(share, [0,1])      &     (0.156)         &                     &                     &                     \\
[1em]
--of~which~face-to-face&                     &       0.107         &       0.343         &       0.400         \\
intensive           &                     &     (0.531)         &     (0.486)         &     (0.760)         \\
[1em]
Presence-intensive~workers&       0.224\sym{**} &       0.079         &      -0.053         &       0.005         \\
(share, [0,1])      &     (0.111)         &     (0.120)         &     (0.128)         &     (0.107)         \\
[1em]
Change~in~number~of &                     &                     &       0.188\sym{***}&      -0.084         \\
monthly visits (log)&                     &                     &     (0.059)         &     (0.122)         \\
[1em]
--times~customer-facing~share&                     &                     &                     &       1.105\sym{***}\\
                    &                     &                     &                     &     (0.379)         \\
[1em]
--times~teamwork-intensive~share&                     &                     &                     &      -0.118         \\
                    &                     &                     &                     &     (1.270)         \\
\hline
Observations        &          79         &          79         &          78         &          78         \\
\(R^{2}\)           &       0.245         &       0.183         &       0.307         &       0.452         \\
\hline\hline
\end{tabular}
}