### 2020-07-01 text/regression.tex
Use April for both CES employment and SafeGraph visits
Results are essentially identical. Resolves #13

{\setlength{\tabcolsep}{-2pt}
{\def\sym#1{\ifmmode^{#1}\else\(^{#1}\)\fi}\begin{tabular}{l*{4}{D{.}{.}{-1}}}\hline\hline
                    &\multicolumn{1}{c}{(1)}&\multicolumn{1}{c}{(2)}&\multicolumn{1}{c}{(3)}&\multicolumn{1}{c}{(4)}\\
                    &\multicolumn{1}{c}{} &\multicolumn{1}{c}{} &\multicolumn{1}{c}{} &\multicolumn{1}{c}{} \\
\hline
Customer-facing~workers&      -0.311\sym{**} &                     &                     &                     \\
\hspace*{1em}(share, [0,1])&     (0.142)         &                     &                     &                     \\
[1em]
\hspace*{1em}of~which~face-to-face&                     &      -0.418\sym{**} &      -0.463\sym{***}&       0.012         \\
\hspace*{1em}intensive (share, [0,1])&                     &     (0.164)         &     (0.152)         &     (0.173)         \\
[1em]
Teamwork-intensive~workers&       0.676\sym{***}&                     &                     &                     \\
\hspace*{1em}(share, [0,1])&     (0.160)         &                     &                     &                     \\
[1em]
\hspace*{1em}of~which~face-to-face&                     &       0.024         &       0.254         &       0.600         \\
\hspace*{1em}intensive (share, [0,1])&                     &     (0.563)         &     (0.532)         &     (0.839)         \\
[1em]
Presence-intensive~workers&       0.218\sym{*}  &       0.079         &      -0.051         &      -0.005         \\
\hspace*{1em}(share, [0,1])&     (0.113)         &     (0.125)         &     (0.136)         &     (0.113)         \\
[1em]
Change~in~number~of &                     &                     &       0.185\sym{***}&      -0.119         \\
\hspace*{1em}monthly visits (log)&                     &                     &     (0.063)         &     (0.131)         \\
[1em]
\hspace*{1em}$\times$~customer-facing~share&                     &                     &                     &       1.021\sym{**} \\
\hspace*{1em}([0,1])&                     &                     &                     &     (0.447)         \\
[1em]
\hspace*{1em}$\times$~teamwork-intensive~share&                     &                     &                     &       0.332         \\
\hspace*{1em}([0,1])&                     &                     &                     &     (1.500)         \\
\hline
Observations        &          79         &          79         &          78         &          78         \\
\(R^{2}\)           &       0.245         &       0.187         &       0.302         &       0.435         \\
\hline\hline
\end{tabular}
}